Of judges, aliens and tpos

### 2020-12-30 paper/paper.tex
fix: fixed figures floating on random pages
feat: conclusion

\documentclass[english, 12pt]{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{dsfont}
\usepackage{enumitem}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage[onehalfspacing]{setspace}

@@

% TODO abstract
\abstract{Total preorders are a common tool to model plausibility orderings over worlds for agents in the research field of belief change. In their paper 'How to Revise a Total Preorder', Booth and Meyer present an approach to revising preorders for iterated belief revision based on assigning abstract intervals of plausibility under new evidence to worlds.

@@

In the field of belief change total preorders (or tpos) are a common tool to model an agent's conditional beliefs. To support iterated belief revision it is necessary to research how these structures can be changed to adapt to an agent's changing world view.
Booth and Meyer suggest an approach based on the idea of assigning an interval of additional metadata to the ordered, propositional worlds and iterate the tpo of an agent using that structure. They define and discuss properties of operators for tpo-revision and relate their approach to other fields of research.
This report on their paper aims to present parts of their work to a wider audience with the addition of an accompanying example. The report's focus will be on the tpo-revision operators discussed in the paper and explicitly avoid the addition of strict preference hierarchies and relations to other research areas.
The structure of the report will be as follows: First it will establish the context of the paper by presenting the research area of iterated belief revision and how it relates to other areas of artificial intelligence and philosophy. To establish a shared understanding of notation and definitions section \ref{section:formal-background} will discuss formal background and other, previous approaches to belief change. It will also introduce the main contribution of this report in an accompanying example demonstrating tpo-revision as described by Booth and Meyer. After establishing a foundation, section \ref{section:booth-and-meyer-tpo-revision-operators} and \ref{section:properties-of-bm-tpo-revision-operators} present part of the work from the original paper 'How to Revise a Total Preorder' \cite{Booth2011}. Section \ref{section:booth-and-meyer-tpo-revision-operators} establishes and defines functions $\ast$, used to revise tpos, a core contribution of Booth and Meyer. In section \ref{section:properties-of-bm-tpo-revision-operators} properties of these functions will be presented. To provide an outlook for the additional work done by Booth and Meyer, \ref{section:iterating-preceq-revision} shows their suggestion for a concrete operator for tpo-revision and includes several examples and novel visualisations of it being applied in the context of the accompanying example. Finally section \ref{section:discussion} closes the report with a summary and commentary on Booth and Meyer's approach to iterated belief revision with interval orderings.

@@

% TODO: discussed in \cite{Katsuno1991} correct? was mentioned in intro of Admissible and Restrained Revision
A common tool used to encode plausibility assumptions of an agent about possible worlds are total preorders (shortened as tpos) \cite{Booth2011}, discussed in \cite{Katsuno1991}. The primary article for this report, 'How to Revise a Total Preorder', researches how to model the change of these total preorders when new evidence becomes available. Being able to derive a new tpo from additional evidence is an approach to handling multiple revision steps and therefor places the article in the field of iterated belief revision.
%1.2 Ich sehe hier die Gefahr, dass das zu Ausführlich wird.
%Keine große Diskussion von AGM, Darwiche and Pearl und so weiter.
%Keine Beziehung zur Allen und co herstellen ;)
%Kurz und Prägnant präsentieren, was benötgigt wird um zu Verstehen was
%in dem Primärartikel neu ist.

@@

    \begin{table}[h]
         \centering
        \begin{tabular}{llll}
        $R_{1}$                      & $R_{2}$                                                                   & $R_{3}$ & $R_{4}$                      \\ \hline
        \multicolumn{1}{|l|}{\begin{tabular}[c]{@{}l@{}}$010$\\ $100$\end{tabular}} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}$000$\\ $110$\end{tabular}} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}$011$\\ $101$\end{tabular}} &
            \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}$001$\\ $111$\end{tabular}} \\ \hline
        \end{tabular}
        \caption{Visualising a tpo as a linearly ordered set of ranks as done in \cite{Booth2006}}
        \label{tab:visualising-a-tpo-example}
    \end{table}
\end{example}

@@

    \begin{figure}[h]
        \centering
        \begin{tikzpicture}[scale=1]
            \begin{axis}[
                  yticklabels={,,$x_{4}$,$x_{3}$,$x_{2}$,$x_{1}$},
                  xticklabels={,,},
                  ytick style={draw=none},
                  xtick style={draw=none},
                  axis line style={draw=none}
            ]
            \addplot[
                scatter,
                scatter src=explicit symbolic,
                mark size=3,
                scatter/classes={
                    empty={mark=*, fill=white},
                    filled={mark=*, fill=black}
                },
                nodes near coords*={\Label},
                visualization depends on={value \thisrow{label} \as \Label}
            ]
            table [meta=class] {
                    x y class label

@@

% TODO do we need this?
\newpage

@@

    \begin{figure}[h]
        \centering
        \begin{tikzpicture}[scale=1]
            \begin{axis}[
                  yticklabels={,,},
                  xticklabels={,,},
                  ytick style={draw=none},
                  xtick style={draw=none},
                  axis line style={draw=none}
            ]
            \addplot[
                scatter,
                scatter src=explicit symbolic,
                mark size=3,
                scatter/classes={
                    empty={mark=*, fill=white},
                    filled={mark=*, fill=black}
                },
                nodes near coords*={\Label},
                visualization depends on={value \thisrow{label} \as \Label}
            ]
            table [meta=class] {
                    x y class label

@@

% TODO: check all \newpages if I need them or fix the figure problem

% TODO do we need this?
\newpage

@@

    \begin{figure}[h]
            \centering
            \begin{tikzpicture}[scale=1.5]
                \begin{axis}[
                        xticklabels={},
                        yticklabels={},
                        extra y ticks={1, 2, 3, 4, 5, 6, 7, 8},
                        extra y tick labels={$111$, $001$, $101$, $011$, $110$, $000$, $100$, $010$},
                        ytick style={draw=none},
                        xtick style={draw=none},
                        axis line style={draw=none}
                ]
                \addplot[
                    scatter,
                    scatter src=explicit symbolic,
                    mark size=3,
                    scatter/classes={
                        empty={mark=*, fill=white},
                        filled={mark=*, fill=black}
                    },
                    nodes near coords*={\Label},
                    visualization depends on={value \thisrow{label} \as \Label}
                ]
                table [meta=class] {
                        x y class label

@@

    Using the $\leq$-faithful tpo $\preceq$ already established for the courtroom example \ref{example:example-introduction} the revision by $\alpha = r$ ("the victim was an alien"). is shown in figure \ref{fig:example-non-prio-revision}.
    \begin{figure}[h]
            \centering
            \begin{tikzpicture}[scale=1.5]
                \begin{axis}[
                        xticklabels={},
                        yticklabels={},
                        extra y ticks={1, 2, 3, 4, 5, 6, 7, 8},
                        extra y tick labels={$111$, $001$, $101$, $011$, $110$, $000$, $100$, $010$},
                        ytick style={draw=none},
                        xtick style={draw=none},
                        axis line style={draw=none}
                ]
                \addplot[
                    scatter,
                    scatter src=explicit symbolic,
                    mark size=3,
                    scatter/classes={
                        empty={mark=*, fill=white},
                        filled={mark=*, fill=black}
                    },
                    nodes near coords*={\Label},
                    visualization depends on={value \thisrow{label} \as \Label}
                ]
                table [meta=class] {
                        x y class label

@@

% TODO write Properties of Booth and Meyer tpo-revision operators
\section{Properties of BM tpo-revision operators}
\label{section:properties-of-bm-tpo-revision-operators}
While not determining a concrete operator $\ast_{\preceq}$ to revise a tpo $\leq$, definition \ref{definition:revision-operator} allows Booth and Meyer to discuss what properties any operator in that family must have \cite{Booth2011}.

@@

With these properties for $\ast$ defined, Booth and Meyer define the family of operators they are discussing with theorem \ref{theorem:revision-operator}.

% THEOREM 1 as axiomatisation of the family of revision operators considered in the paper. "It turns out that these properties provide an exact characterisation of the
% revision operators we consider."

@@

    \begin{figure}[h]
            \centering
            \begin{tikzpicture}[scale=1.5]
                \begin{axis}[
                        xticklabels={},
                        yticklabels={},
                        extra y ticks={1, 2, 3, 4, 5, 6, 7, 8},
                        extra y tick labels={$111$, $001$, $101$, $011$, $110$, $000$, $100$, $010$},
                        ytick style={draw=none},
                        xtick style={draw=none},
                        axis line style={draw=none}
                ]
                \addplot[
                    scatter,
                    scatter src=explicit symbolic,
                    mark size=3,
                    scatter/classes={
                        empty={mark=*, fill=white},
                        old={mark=x, fill=white}
                    },
                    nodes near coords*={\Label},
                    visualization depends on={value \thisrow{label} \as \Label}
                ]
                table [meta=class] {
                        x y class label

@@

% TODO need this?
\newpage

@@

    \begin{figure}[h]
            \centering
            \begin{tikzpicture}[scale=1.5]
                \begin{axis}[
                        xticklabels={},
                        yticklabels={},
                        extra y ticks={1, 2, 3, 4, 5, 6, 7, 8},
                        extra y tick labels={$111$, $001$, $101$, $011$, $110$, $000$, $100$, $010$},
                        ytick style={draw=none},
                        xtick style={draw=none},
                        axis line style={draw=none}
                ]
                \addplot[
                    scatter,
                    scatter src=explicit symbolic,
                    mark size=3,
                    scatter/classes={
                        empty={mark=*, fill=white},
                        filled={mark=*, fill=black},
                        old={mark=x, fill=white}
                    },
                    nodes near coords*={\Label},
                    visualization depends on={value \thisrow{label} \as \Label}
                ]
                table [meta=class] {
                        x y class label

@@

For the initial tpo $\leq$ there was no possible evidence that would change the plausibility ordering between worlds where the victim is an alien and worlds where they are not. That was reflected by the new tpo $\leq_{\alpha}^{\ast}$ that even with direct evidence for alien life did not change the judges plausibility orderings. But because of the revision of $\preceq$ to $\preceq_{p \ast \alpha}$ the evidence changed the acceptance of the judge to belief we are not alone in the universe. After receiving $\beta$ the judge even considers the $[r]$-worlds $\{011, 101\}$ more likely than the $[\neg r]$-worlds $\{000, 110\}$!

@@

% TODO write Discussion
\section{Discussion}
\label{section:discussion}
% TODO *6,*7 seem weird, no different "strength" of evidence? maybe revising multiple time by the same evidence should that be? otherwise revising multiple times by same evidence is also weird because it changes things and is not idempotent
% TODO: discuss 6/7 being wrong? strength of evidence?=!=!, discuss 5 with it being equal when sticks - and + are same rank?! 
\newpage

>>>>

\documentclass[english, 12pt]{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{dsfont}
\usepackage{enumitem}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage[onehalfspacing]{setspace}
\usepackage{float}

@@

\abstract{Total preorders are a common tool to model plausibility orderings over worlds for agents in the research field of belief change. In their paper 'How to Revise a Total Preorder', Booth and Meyer present an approach to revising preorders for iterated belief revision based on assigning abstract intervals of plausibility under new evidence to worlds.

@@

In the field of belief change total preorders (or tpos) are a common tool to model an agent's conditional beliefs. To support iterated belief revision it is necessary to research how these structures can be changed to adapt to an agent's changing world view.
Booth and Meyer suggest an approach based on the idea of assigning an interval of additional metadata to the ordered, propositional worlds and iterate the tpo of an agent using that structure. They define and discuss properties of operators for tpo-revision and relate their approach to other fields of research.
This report on their paper aims to present parts of their work to a wider audience with the addition of an accompanying example. The report's focus will be on the tpo-revision operators discussed in the paper and explicitly avoid the addition of strict preference hierarchies and relations to other research areas.
The structure of the report will be as follows: First it will establish the context of the paper by presenting the research area of iterated belief revision and how it relates to other areas of artificial intelligence and philosophy. To establish a shared understanding of notation and definitions section \ref{section:formal-background} will discuss formal background and other, previous approaches to belief change. It will also introduce the main contribution of this report in an accompanying example demonstrating tpo-revision as described by Booth and Meyer. After establishing a foundation, section \ref{section:booth-and-meyer-tpo-revision-operators} and \ref{section:properties-of-bm-tpo-revision-operators} present part of the work from the original paper 'How to Revise a Total Preorder' \cite{Booth2011}. Section \ref{section:booth-and-meyer-tpo-revision-operators} establishes and defines functions $\ast$, used to revise tpos, a core contribution of Booth and Meyer. In section \ref{section:properties-of-bm-tpo-revision-operators} properties of these functions will be presented. To provide an outlook for the additional work done by Booth and Meyer, \ref{section:iterating-preceq-revision} shows their suggestion for a concrete operator for tpo-revision and includes several examples and adapted visualisations of it being applied in the context of the accompanying example. Finally section \ref{section:conclusion} closes the report with a summary and commentary on Booth and Meyer's approach to iterated belief revision with interval orderings.

@@

% TODO: discussed in \cite{Katsuno1991} correct? was mentioned in intro of Admissible and Restrained Revision
A common tool used to encode plausibility assumptions of an agent about possible worlds are total preorders (shortened as tpos) \cite{Booth2011}, discussed in \cite{Katsuno1991}. The primary article for this report, 'How to Revise a Total Preorder', researches how to model the change of these total preorders when new evidence becomes available. Being able to derive a new tpo from additional evidence is an approach to handling multiple revision steps and therefor places the article in the field of iterated belief revision.

@@

    \begin{table}[H]
         \centering
        \begin{tabular}{llll}
        $R_{1}$                      & $R_{2}$                                                                   & $R_{3}$ & $R_{4}$                      \\ \hline
        \multicolumn{1}{|l|}{\begin{tabular}[c]{@{}l@{}}$010$\\ $100$\end{tabular}} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}$000$\\ $110$\end{tabular}} & \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}$011$\\ $101$\end{tabular}} &
            \multicolumn{1}{l|}{\begin{tabular}[c]{@{}l@{}}$001$\\ $111$\end{tabular}} \\ \hline
        \end{tabular}
        \caption{Visualising a tpo as a linearly ordered set of ranks as done in \cite{Booth2006}}
        \label{tab:visualising-a-tpo-example}
    \end{table}
\end{example}

@@

    \begin{figure}[H]
        \centering
        \begin{tikzpicture}[scale=1.5]
            \begin{axis}[
                  yticklabels={,,$x_{4}$,$x_{3}$,$x_{2}$,$x_{1}$},
                  xticklabels={,,},
                  ytick style={draw=none},
                  xtick style={draw=none},
                  axis line style={draw=none}
            ]
            \addplot[
                scatter,
                scatter src=explicit symbolic,
                mark size=3,
                scatter/classes={
                    empty={mark=*, fill=white},
                    filled={mark=*, fill=black}
                },
                nodes near coords*={\Label},
                visualization depends on={value \thisrow{label} \as \Label}
            ]
            table [meta=class] {
                    x y class label

@@

@@

    \begin{figure}[H]
        \centering
        \begin{tikzpicture}[scale=1.5]
            \begin{axis}[
                  yticklabels={,,},
                  xticklabels={,,},
                  ytick style={draw=none},
                  xtick style={draw=none},
                  axis line style={draw=none}
            ]
            \addplot[
                scatter,
                scatter src=explicit symbolic,
                mark size=3,
                scatter/classes={
                    empty={mark=*, fill=white},
                    filled={mark=*, fill=black}
                },
                nodes near coords*={\Label},
                visualization depends on={value \thisrow{label} \as \Label}
            ]
            table [meta=class] {
                    x y class label

@@

@@

    \begin{figure}[H]
            \centering
            \begin{tikzpicture}[scale=1.5]
                \begin{axis}[
                        xticklabels={},
                        yticklabels={},
                        extra y ticks={1, 2, 3, 4, 5, 6, 7, 8},
                        extra y tick labels={$111$, $001$, $101$, $011$, $110$, $000$, $100$, $010$},
                        ytick style={draw=none},
                        xtick style={draw=none},
                        axis line style={draw=none}
                ]
                \addplot[
                    scatter,
                    scatter src=explicit symbolic,
                    mark size=3,
                    scatter/classes={
                        empty={mark=*, fill=white},
                        filled={mark=*, fill=black}
                    },
                    nodes near coords*={\Label},
                    visualization depends on={value \thisrow{label} \as \Label}
                ]
                table [meta=class] {
                        x y class label

@@

    Using the $\leq$-faithful tpo $\preceq$ already established for the courtroom example \ref{example:example-introduction} the revision by $\alpha = r$ ("the victim was an alien"). is shown in figure \ref{fig:example-non-prio-revision}.
    \begin{figure}[H]
            \centering
            \begin{tikzpicture}[scale=1.5]
                \begin{axis}[
                        xticklabels={},
                        yticklabels={},
                        extra y ticks={1, 2, 3, 4, 5, 6, 7, 8},
                        extra y tick labels={$111$, $001$, $101$, $011$, $110$, $000$, $100$, $010$},
                        ytick style={draw=none},
                        xtick style={draw=none},
                        axis line style={draw=none}
                ]
                \addplot[
                    scatter,
                    scatter src=explicit symbolic,
                    mark size=3,
                    scatter/classes={
                        empty={mark=*, fill=white},
                        filled={mark=*, fill=black}
                    },
                    nodes near coords*={\Label},
                    visualization depends on={value \thisrow{label} \as \Label}
                ]
                table [meta=class] {
                        x y class label

@@

\section{Properties of BM tpo-revision operators}
\label{section:properties-of-bm-tpo-revision-operators}
While not determining a concrete operator $\ast_{\preceq}$ to revise a tpo $\leq$, definition \ref{definition:revision-operator} allows Booth and Meyer to discuss what properties any operator in that family must have \cite{Booth2011}.

@@

With these properties for $\ast$ discussed, Booth and Meyer define the family of operators they are discussing with theorem \ref{theorem:revision-operator}.

@@

    \begin{figure}[H]
            \centering
            \begin{tikzpicture}[scale=1.5]
                \begin{axis}[
                        xticklabels={},
                        yticklabels={},
                        extra y ticks={1, 2, 3, 4, 5, 6, 7, 8},
                        extra y tick labels={$111$, $001$, $101$, $011$, $110$, $000$, $100$, $010$},
                        ytick style={draw=none},
                        xtick style={draw=none},
                        axis line style={draw=none}
                ]
                \addplot[
                    scatter,
                    scatter src=explicit symbolic,
                    mark size=3,
                    scatter/classes={
                        empty={mark=*, fill=white},
                        old={mark=x, fill=white}
                    },
                    nodes near coords*={\Label},
                    visualization depends on={value \thisrow{label} \as \Label}
                ]
                table [meta=class] {
                        x y class label

@@

@@

    \begin{figure}[H]
            \centering
            \begin{tikzpicture}[scale=1.5]
                \begin{axis}[
                        xticklabels={},
                        yticklabels={},
                        extra y ticks={1, 2, 3, 4, 5, 6, 7, 8},
                        extra y tick labels={$111$, $001$, $101$, $011$, $110$, $000$, $100$, $010$},
                        ytick style={draw=none},
                        xtick style={draw=none},
                        axis line style={draw=none}
                ]
                \addplot[
                    scatter,
                    scatter src=explicit symbolic,
                    mark size=3,
                    scatter/classes={
                        empty={mark=*, fill=white},
                        filled={mark=*, fill=black},
                        old={mark=x, fill=white}
                    },
                    nodes near coords*={\Label},
                    visualization depends on={value \thisrow{label} \as \Label}
                ]
                table [meta=class] {
                        x y class label

@@

For the initial tpo $\leq$ there was no possible evidence that would change the plausibility ordering between worlds where the victim is an alien and worlds where they are not. That was reflected by the new tpo $\leq_{\alpha}^{\ast}$ that even with direct evidence for alien life did not change the judges plausibility orderings. But because of the revision of $\preceq$ to $\preceq_{p \ast \alpha}$, the evidence $\alpha$ changed the acceptance of the judge to belief we are not alone in the universe. After receiving $\beta$ the judge even considers the $[r]$-worlds $\{011, 101\}$ more likely than the $[\neg r]$-worlds $\{000, 110\}$!

@@

% TODO write Conclusion
\section{Conclusion}
\label{section:conclusion}
The featured paper 'How to Revise a Total Preorder' by Booth and Meyer \cite{Booth2011} is an expansive work on using interval orderings for belief change. This report focused on presenting the research context and previous work in iterated belief change as well as the concrete tpo-revision based on the additional structure of $\leq$-faithful tpos. A consistent example based on a trial situation (adapted from \cite{Darwiche1997} and \cite{Booth2011}) was used to accompany the discussion of the paper. In addition to the provided example the report also introduced an adapted visualisation for $\leq$-faithful tpo-revision based on the stick visualisation used by Booth and Meyer themselves.

To make the presented content easier to follow some parts of the original paper had to be excluded. Booth and Meyer discuss multiple additional properties of their proposed operators as well as extreme examples. They also provide a different way to express the $\preceq$-structure using strict preference hierarchies and discuss how they relate to other research areas. For those topics as well as detailed proofs it is highly advised to read the primary paper \cite{Booth2011} itself.

The idea to define an interval of plausibility for worlds depending on if evidence supports them or not does make intuitive sense. An angle largely missing from the paper is the discussion of the different impact and quality of new information itself.

% idempotent correct?
The proposed concrete operator for the revision of $\leq$-faithful tpos allows for multiple revisions by the same evidence $\alpha$ or semantically equivalent evidence $\beta \equiv \alpha$, consistently lowering the acceptance of $[\neg\alpha]$-worlds without presenting new evidence. It seems like an interesting question to ask if tpo-revision for iterated belief change should be idempotent in regards to already applied evidence.

Considering quality and impact of new evidence also creates the question if \ref{AST6} and \ref{AST7} (properties of BM tpo-revision operators, page \ref{section:properties-of-bm-tpo-revision-operators}) should be weakened. Especially high quality evidence for a very unlikely type of world might change an agents plausibility ordering where low quality evidence that is very expected might not. Consider the following situation in our courtroom example \ref{example:example-introduction}: New evidence $\alpha$ arrives that fingerprints of John have been found at the place of the murder, pointing to a world $x$ with $p$ being true. But, because of a large body of evidence, the judge still considers a world $y$ where Mary is the murderer to be more plausible. This means there can be no evidence $\gamma$ that would make the agent consider the world $x$ more likely than $y$ according to \ref{AST7}:

\begin{equation*}
    \textrm{If } x \in [\alpha], y \in [\neg\alpha] \textrm{ and } y <_{\alpha}^{\ast} x \textrm{ then } y <_{\gamma}^{\ast} x
\end{equation*}

But what if evidence is submitted that if Mary is the murderer, the victim must be an alien $\gamma = q \implies r$? Semantically a human agent would assume all worlds in which Mary is the murderer to be highly unlikely due to how far fetched the acceptance of the victim being an alien is. How to solve this apparent contradiction by modelling different types and strengths of evidence could be a future area of research.

% TODO *6,*7 seem weird, no different "strength" of evidence? maybe revising multiple time by the same evidence should that be? otherwise revising multiple times by same evidence is also weird because it changes things and is not idempotent, discuss 5 with it being equal when sticks - and + are same rank?! 
\newpage
